---
title: 'Homework #8'
author: "Andrea Astudillo"
date: "2024-11-23"
output: html_document
---

```{r}
library(fpp)
library(fpp2)
library(TTR)
library(forecast)

Tesla.Model.Y <- read.csv(file.choose())
salesY=(Tesla.Model.Y)

ts_salesY = ts(salesY$Sales.Units,start = c(2020,7), frequency = 12)
plot(ts_salesY)

nsdiffs(ts_salesY)
ndiffs(ts_salesY)
ndiffs((diff(ts_salesY,12)))
tsdisplay(diff(diff(ts_salesY,12)))
tsdisplay(ts_salesY)
fit3 <- auto.arima(ts_salesY,trace=TRUE, stepwise = FALSE )
fit3
res = residuals(fit3)
tsdisplay(res)
```
In the tsdisplay plot we can observe that there some spikes. In the ACF we can see there are spikes in lag 2 to lag 8 and in PACF there is a spike in lag 2.

In tsdisplay(res) plot we can now observe that there are no spikes in neither ACF and PACF.

sar1=0.6054 indicates that the seasonal values have a strong effect on the current values

s.e=0.1535 is the standard error indicating that the model is reliable for estimates.

```{r}
#Residuals
Acf(fit3$residuals)
Box.test(residuals(fit3), lag=20, type="Ljung")
plot.ts(residuals(fit3))
hist(fit3$residuals)
```

In ACF plot we can now observe that there are no spikes. It can also be confirmed with the plot which the residuals have a random pattern. The same for the histogram, majority of the values are at zero indicating it is normally distributed.



```{r}
#Forecast
plot(forecast(fit3, h=5))
lines(fit3$fitted,col="red")
accuracy(forecast(fit3))
```
In the accuracy forecast we can observe that ME has the lowest accuracy result.