---
title: "Untitled"
author: "Andrea Astudillo"
date: "2024-11-18"
output: html_document
---

```{r}
library(fpp)
library(fpp2)
library(TTR)
library(UsingR)
library(ggplot2)

Tesla.Model.Y <- read.csv(file.choose())
salesY=(Tesla.Model.Y)

plot(salesY$Sales.Units, salesY$Gas.Price, main = "Scatterplot")

#1-Perform regression
plot(salesY)
summary(salesY)
cor(salesY$Sales.Units, salesY$Gas.Price)

fit = lm(salesY$Sales.Units ~ salesY$Gas.Price, data = salesY)
ggplot(salesY, aes(x=salesY$Sales.Units, y=salesY$Gas.Price)) +geom_point() + geom_smooth(method = "lm")
fit
summary(fit)
anova(fit)

fit2 = lm(log(Sales.Units) ~ log(Gas.Price), data = salesY)
plot(jitter(Sales.Units)~jitter(Gas.Price),xlab="Sales Units", ylab = "Gas Price", data = salesY)
lines(1:40, exp(fit2$coef[1]+fit2$coef[2]*log(1:40)))

plot(log(jitter(salesY$Sales.Units)) ~ log(jitter(salesY$Gas.Price)),xlab="log Sales Units0", ylab="lab Gas Price", data = salesY)
abline(fit2)
```

```{r}
#2-Perform Model Acceptance Steps and show the interpretation of the model output
summary(fit2)
```

The model appears to be statistically significant.
-43% of the variance of the sales units of model Y can be explained by the variation of the gas prices
-Also the F-statistic is 32.9 indicating there is a linear regression between the gas prices and the sales value. 
-As well for the p value with 1.11e-06, indicating is far away from the center space of the confidence interval.

```{r}
#3-Perform Residual Analysis and explain the insights (3 Points)
res = residuals(fit2)
par(mfrow=c(1,2))
plot(jitter(res, amount=0.005) ~ jitter(log(Gas.Price)),ylab = "Residuals", xlab = "Log Gas Price", data=salesY)
```

Ideally the residual plot should have not pattern where the gas price will have no effect on the residuals. However, the plot above we have observe some pattern.

